---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r , echo = TRUE}
if(!file.exists("activity.csv")){
  unzip("activity.zip")
}
activity_data <- read.csv("activity.csv")
activity_data <- activity_data[!is.na(activity_data[,"steps"]),]
activity_data$date <- as.Date(as.character(activity_data$date))

```


## What is mean total number of steps taken per day?

```{r, echo = TRUE}
Grouped_data <- aggregate(activity_data$steps, by = list(activity_data$date), sum)
names(Grouped_data) <- c("date", "total_steps")

hist(Grouped_data$total_steps, xlab = "Steps", main = "Total Daily Steps", col = "steelblue")
print(paste("Mean number of steps taken daily", mean(Grouped_data$total_steps)))
print(paste("Median number of steps taken daily", median(Grouped_data$total_steps)))
```

## What is the average daily activity pattern?

```{r, echo = TRUE}
library(ggplot2)
Mean_Data_By_Interval <- aggregate(activity_data$steps, by=list(activity_data$interval), mean)
names(Mean_Data_By_Interval) <- c("interval", "steps")

ggplot(Mean_Data_By_Interval, aes(x = interval, y=steps)) +
  labs(title = "Sum of Steps by Interval", x = "interval", y = "steps")+
  geom_line(color="red") 

maxInterval <- Mean_Data_By_Interval[which.max(Mean_Data_By_Interval$steps),]
maxInterval
```


## Imputing missing values

```{r}
activity_data <- read.csv("activity.csv")
missing_values <- is.na(activity_data$steps)
print(paste("Number of Missing values are", sum(missing_values)))

clean_activity_data <- activity_data[!missing_values,]
mean_values <- tapply(clean_activity_data$steps, clean_activity_data$interval, mean, na.rm=TRUE, simplify=TRUE)
activity_data$steps[missing_values] <- mean_values[as.character(activity_data$interval[missing_values])]
print(paste("Number of Missing values after imputing", sum(is.na(activity_data$steps))))

Grouped_data <- aggregate(activity_data$steps, by = list(activity_data$date), sum)
names(Grouped_data) <- c("Date", "total_steps")
hist(Grouped_data$total_steps, xlab = "Steps", main = "Total Daily Steps", col = "steelblue")
print(paste("Mean number of steps taken daily", mean(Grouped_data$total_steps)))
print(paste("Median number of steps taken daily", median(Grouped_data$total_steps)))
```

The original mean and median were 10766.1886792453 and 10765, respectively.
Mean value remained the same while median has increased by 1.1886792453.

The effect of using mean data per interval as a data impute method for missing values seems to push overall data towards the mean.

## Are there differences in activity patterns between weekdays and weekends?
